### 2014-09-29 PrivacyProtection-2014-09-09.tex
Resolved conflict

\chapter{Legal and Ethical aspects of Privacy}

@@

\section{Privacy Definition and Taxonomy}

@@

At first, one has to outline the current state of the system under study. This includes description of:

\begin{itemize}
\item all \textbf{Assets} which must be protected
\item all relevant \textbf{IT-Systems}
\item all involved \textbf{Humans} and their \textbf{Conflicts of Interests}
\item all known \textbf{Vulnerabilities}
\item and all important \textbf{Interactions} between the former components.
\end{itemize}

@@

We focus on the Asset privacy as described in Section \ref{sec:taxonomy}

@@

@@

A \textit{Citizen (a)} carries a mobile device running the \textit{Sensor Collector (d)} application.
The \textit{Sensor Collector} application collects sensor data (accelerometer, GPS, GSM, ...) and sends it to the \textit{Data Cetner (c)}. 
The \textit{Data Center} on the other side sends beneficial information (traffic jam reports, bus schedule) for \textit{Citizens} to the mobile device.
The \textit{System Provider (b)} provides and operates technical infrastructure like the Data Center and the \textit{Report Tool (f)}.
The \textit{Report Tool} queries the \textit{Data Center} for aggregated data to visualize in form of charts and other means suitable to help understanding of monitored \textit{Citizens}. 
\textit{Local Authorities (c)} use the \textit{Report Tool} to get information in order to understand citizen movement and improve public services.

In such systems, the most valuable information is not the raw sensor data, the beneficial information for \textit{Local Authorities} is produced via \textit{Data Mining} (Human Activity Recognition: detect running, walking, standing, ...). 
Usually, \textit{Data Mining} is heavily conduct within the \textit{Data Center}, but simpler mining processes can also be conducted on mobile devices.

@@

\textbf{Citizen:}
\begin{itemize}
\item legal right to privacy
\item no spam
\item no stalking
\item no loss / gain of money
\item physical wellbeing / personal health (attack because someone knows your gps location)
\item transparency: who has my data
\item good pricing of insurance policies
\end{itemize}

@@

\textbf{Local Authority:}

In short, Local Authorities are interested in ...
\begin{itemize}
\item ... business intelligence (profit, RoI)
\item ... political reputation
\end{itemize}

>>>>

\chapter{Legal and Ethical aspects of Privacy}
\label{chap:privacy}

@@

\section{Privacy Definition and Taxonomy}
\label{sec:taxonomy}

@@

At first, one has to outline the current state of the system under study. This includes description of:
\begin{itemize}
\item all \textbf{Assets} which must be protected
\item all relevant \textbf{IT-Systems}
\item all involved \textbf{Humans} and their \textbf{Conflicts of Interests}
\item all known \textbf{Vulnerabilities}
% \item and all important \textbf{Interactions} between the former components.
\end{itemize}

@@

In this document we focus our attention to only one asset: The privacy of the citizen.

%TODO: UPDATE
Our definition of privacy is descibed in detail in Chapter \ref{chap:privacy}. In Section \ref{sec:taxonomy} we define privacy as the ``control over private data'' and introduce the following seven different types of privacy:
\begin{enumerate}
\item Privacy of the Person
\item Privacy of Behaviour and Action
\item Privacy of Communication
\item Privacy of Data and Image
\item Privacy of Thoughts and Feelings
\item Privacy of Location and Space
\item Privacy os Association
\end{enumerate}

Please refer to Section \ref{sec:taxonomy} for more details.

@@

The IT infrastructure of the Live+Gov system, i.e. the Live+Gov toolkit and the customization of the software components are described in detail in various project deliverables: D4.1, D4.3, D1.1, D5.1.
In this section we give an abstraction of those systems from the perspective of WP1.

@@

A citizen (a) carries a mobile device running the \emph{sensor collector} (d) application.
The sensor collector application collects sensor data (accelerometer, GPS, GSM, ...) and sends it to the \emph{data center} (c).
The sensor collector can also be able to perform certain data mining operations.
Examples of such data mining operations include human activity recognition,the detection of service lines, detection of characters and faces from images.

The \emph{data center} stores and processes sensor data collected with the sensor collector application.
Again the amount

 sends beneficial information (traffic jam reports, bus schedule) for \textit{Citizens} to the mobile device.
The \textit{System Provider (b)} provides and operates technical infrastructure like the \textit{Data Center} and the \textit{Report Tool (f)}.
The \textit{Report Tool} queries the \textit{Data Center} for aggregated data to visualize in form of charts and other means suitable to help understanding of monitored \textit{Citizens}. 
\textit{Local Authorities (c)} use the \textit{Report Tool} to get information in order to understand citizen movement and improve public services.

In such systems, the most valuable information is not the raw sensor data, the beneficial information for \textit{Local Authorities} is produced via \textit{Data Mining} (Human Activity Recognition: detect running, walking, standing, ...). 
Usually, \textit{Data Mining} is heavily conducted within the \textit{Data Center}, but simpler mining processes can also be conducted on mobile devices.

@@

\textbf{Citizen:} 
Citizens are persons who take part in the monitoring system as the monitored subjects.
They offer their privacy sensitive data, because by participating they expect a higher level of comfort in their lives through the offered services in return.
E.g. they will have access to real time bus schedules and traffic jam reports, localized social networking or they generally benefit from public improvement by Local Authorities.

However, they have a legal right to privacy and are interested in having that privacy respected by others.
For example, all humans have a natural interest in physical wellbeing and personal health. 
If someone monitors their immediate location (GPS) or frequently used routes (stalking), people with access to that information could attack and harm citizens.

In general, citizens are not interested in having information about them exploited by others. 
This includes just annoying spam, where their technical identity is used to send them unwanted commercials.
But it also includes more severe phishing attacks, which try to rob citizens of their TAN numbers and disrespect their financial interests.

Such financial interests of monitored citizens, the interest in not losing / gaining money, could be disrespected on an even larger scale.
Collected data could be used for scoring and analysis, which may affect the pricing of policies or credits, if sold to insurances or banks.

So citizens have an inherent interest in transparency and disclosure of who uses their data for which purposes, as it may affect their physical and financial wellbeing.

@@

\textbf{Local Authority:} 
Local Authorities are public offices (ministry, agency, department, ...) or other external public service providers which act as direct customers of Service Providers.
They order a proposed monitoring and mining system specialized for their needs. 
For example a low level department, like departments for public mobility, orders a monitoring and mining system to better understand and improve urban traffic flow.

Such systems are heavy investments, so naturally Local Authorities are interested in a profitable return.
However, the return of investment is not directly of financial nature.
Like Service Providers their financial gain depends on customer satisfaction. 
Customers for Local Authorities are either citizens, who use their services, or politicians, who order their services.
The happiness of both sides is interdependent. 
Citizens are happy customers if the services, e.g. public mobility, work well. 
If citizens are happy, it is more likely that politicians gain reputation, as they organize the public services through Local Authorities.

One way of achieving these business objectives is by utilizing business intelligence. 
A proposed monitoring and mining system can provide such intelligence.
For public mobility services it can provide traffic jam detection to maximize traffic flow.
Or it can provide user profiling to maximize (targeted) publicity.

Additionally, since Local Authorities act like corporations comparable to Service Providers, they are also interested in good working conditions.
As discontent employees may harm the system by disclosure of business intelligence.

In short, Local Authorities are interested in ...
\begin{itemize}
\item ... business intelligence
\item ... political reputation
\item ... good working conditions
\end{itemize}
